---
title: "Assignment 1"
output: html_document
date: "2024-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the packages and data

Loading in the packages:

```{r setup}
library(tidyverse)
library(ggplot2)
library(here)
library(dplyr)
```

Now, we will load in the Eno data and make separate dataframes for the median 
outflow, mean outflow, and volume.

```{r Eno data}
#Eno data:
eno_outflows <- read.csv(
  file = here("./Data Raw/Eno_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median outflow:
eno_average_median <- eno_outflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean outflow:
eno_average_mean <- eno_outflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
eno_volume_mean <- eno_outflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

Now, we will load in the Flat data and make separate dataframes for the median 
outflow, mean outflow, and volume.

```{r Flats data}
#Flats data:
flat_outflows <- read.csv(
  file = here("./Data Raw/Flat_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median outflow:
flat_average_median <- flat_outflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean outflow:
flat_average_mean <- flat_outflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
flat_volume_mean <- flat_outflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

Now, we will load in the Neuse data and make separate dataframes for the median 
inflow, mean inflow, and volume.

```{r Neuse data}
#Neuse data:
neuse_inflows <- read.csv(
  file = here("./Data Raw/Falls_Lake_outlet_Neuse_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median inflow:
neuse_average_median <- neuse_inflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean inflow:
neuse_average_mean <- neuse_inflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
neuse_volume_mean <- neuse_inflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

## Plots



```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
