---
title: 'Assignment 1: Aidan Power, Fiona Price, and Emily Burghart'
output:
  pdf_document: default
  html_document: default
date: "2024-09-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the packages and data

Loading in the packages:

```{r packages}
library(tidyverse)
library(ggplot2)
library(here)
library(dplyr)
library(ggpmisc)
library(extrafont)
```

Now, we will load in the Eno data and make separate dataframes for the median 
inflow, mean inflow, and volume.

```{r Eno data}
#Eno data:
eno_inflows <- read.csv(
  file = here("./Data Raw/Eno_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median inflow:
eno_average_median <- eno_inflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean outflow:
eno_average_mean <- eno_inflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
eno_volume_mean <- eno_inflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

Now, we will load in the Flat data and make separate dataframes for the median 
outflow, mean outflow, and volume.

```{r Flats data}
#Flats data:
flat_inflows <- read.csv(
  file = here("./Data Raw/Flat_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median inflow:
flat_average_median <- flat_inflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean inflow:
flat_average_mean <- flat_inflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
flat_volume_mean <- flat_inflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

Now, we will load in the Neuse data and make separate dataframes for the median 
outflow, mean outflow, and volume.

```{r Neuse data}
#Neuse data:
neuse_outflows <- read.csv(
  file = here("./Data Raw/Falls_Lake_outlet_Neuse_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median outflow:
neuse_average_median <- neuse_outflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean outflow:
neuse_average_mean <- neuse_outflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
neuse_volume_mean <- neuse_outflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

## Plots
Now, we will start making graphs. First, we will look at Eno. We will start with
all years.

```{r Eno all, echo=FALSE}
#Median Eno inflow line graph
ggplot(data = eno_average_median, aes(x = year, y = avg_median_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE, 
               size = 2.5,
               hjust = -1.7, vjust = 4,
               family = "Times New Roman") +
  labs(title = "Median Eno Inflow", x = "Year", y = "Median Inflow (cfs)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Median Eno inflow scatter plot
eno_med_model <- lm(year ~ avg_median_flow, data = eno_average_median)
r_squared_1 <- summary(eno_med_model)$r.squared

ggplot(data = eno_average_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Eno Outflow", x = "Year", y = "Median Inflow (cfs)") +
  annotate("text", x = 1975, y = 3542400, 
           label = paste("R² =", round(r_squared_1, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)


#Mean Eno inflow line graph
ggplot(data = eno_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  annotate("text", x = 1973, y = 9000000, label = "-9.25 * 10^3 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Mean Eno Inflow", x = "Year", y = "Mean Inflow (cfs)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Eno inflow scatter plot
eno_mean_model <- lm(year ~ avg_mean_flow, data = eno_average_mean)
r_squared_2 <- summary(eno_mean_model)$r.squared

ggplot(data = eno_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Inflow", x = "Year", y = "Mean Inflow (cfs)") +
  annotate("text", x = 1975, y = 3542400, 
           label = paste("R² =", round(r_squared_2, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean volume line graph
ggplot(data = eno_volume_mean, aes(x = year, y = avg_vol)) +
  geom_line() +
  geom_smooth(method = "lm") +
    annotate("text", x = 1973, y = 270000000, label = "-2.87 * 10^5 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Average Eno Volume", x = "Year", y = "Mean Volume (cf)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean volume scatter plot
eno_vol_model <- lm(year ~ avg_vol, data = eno_volume_mean)
r_squared_3 <- summary(eno_vol_model)$r.squared

ggplot(data = eno_volume_mean, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Average Eno Volume", x = "Year", y = "Mean Volume (cf)") +
    annotate("text", x = 1975, y = 3542400, 
           label = paste("R² =", round(r_squared_3, 2)), 
           size = 5) +
   stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)
```

The dam was built in 1981. We will only look at data from 1981 and beyond now.

```{r Eno 1981 and beyond, echo=TRUE}
#Filter out the data to only include 1981 and beyond.
eno_filtered <- eno_inflows %>% 
  filter(year >= 1981)
#Filter out the average median inflow data to only include 1981 and beyond.
eno_avemedian_filtered <- eno_average_median %>% 
  filter(year >= 1981)
#Filter out the average mean inflow data to only include 1981 and beyond.
eno_avemean_filtered <- eno_average_mean %>% 
  filter(year >= 1981)
#Filter out the average volume data to only include 1981 and beyond.
eno_vol_filtered <- eno_volume_mean %>% 
  filter(year >= 1981)

#Median Eno inflow line graph
ggplot(data = eno_avemedian_filtered, aes(x = year, y = avg_median_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) +
  labs(title = "Median Eno Inflow Since 1981", x = "Year", y = "Median Inflow (cfs)")

#Median Eno inflow scatter plot
eno_med_model_1981 <- lm(year ~ avg_median_flow, data = eno_average_median)
r_squared_4 <- summary(eno_med_model_1981)$r.squared

ggplot(data = eno_avemedian_filtered, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Eno Inflow since 1981", x = "Year", y = "Median Inflow (cfs)") +
  annotate("text", x = 2000, y = 3542400, 
           label = paste("R² =", round(r_squared_4, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Eno inflow line graph 
ggplot(data = eno_avemean_filtered, aes(x = year, y = avg_mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
    annotate("text", x = 2010, y = 9000000, label = "4.16 * 10^6 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Mean Eno Inflow Since 1981", x = "Year", y = "Mean Inflow (cfs)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Eno inflow scatter plot
eno_mean_model_1981 <- lm(year ~ avg_mean_flow, data = eno_average_mean)
r_squared_5 <- summary(eno_mean_model_1981)$r.squared

ggplot(data = eno_avemean_filtered, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Inflow Since 1981", x = "Year", y = "Mean Inflow (cfs)") +
   annotate("text", x = 2000, y = 3542400, 
           label = paste("R² =", round(r_squared_5, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Eno volume line graph
ggplot(data = eno_vol_filtered, aes(x = year, y = avg_vol)) +
  geom_line() +
  geom_smooth(method = "lm") +
      annotate("text", x = 2010, y = 270000000, label = "-1.88 * 10^3 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Average Eno Volume Since 1981", x = "Year", y = "Mean Volume (cf)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman")) 

#Mean Eno volume scatter plot
eno_vol_model_1981 <- lm(year ~ avg_vol, data = eno_vol_filtered)
r_squared_6 <- summary(eno_vol_model_1981)$r.squared

ggplot(data = eno_vol_filtered, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Average Eno Volume Since 1981", x = "Year", y = "Mean Volume (cf)") +
     annotate("text", x = 2000, y = 3542400, 
           label = paste("R² =", round(r_squared_6, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) 
```

Now, we will look at Flat data.
```{r Flat all, echo=TRUE}
#Median Flat inflow line graph
ggplot(data = flat_average_median, aes(x = year, y = avg_median_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) +
  labs(title = "Median Flat Inflow", x = "Year", y = "Median Inflow (cfs)")

#Median Flat inflow scatter plot
flat_med_model <- lm(year ~ avg_median_flow, data = flat_average_median)
flat_r <- summary(flat_med_model)$r.squared

ggplot(data = flat_average_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Flat Inflow", x = "Year", y = "Median Inflow (cfs)") +
  annotate("text", x = 1975, y = 3542400, 
           label = paste("R² =", round(flat_r, 2)), 
           size = 5,) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Flat inflow line graph
ggplot(data = flat_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
      annotate("text", x = 1950, y = 21000000, label = "-1.85 * 10^3 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Mean Flat Inflow", x = "Year", y = "Mean Inflow (cfs)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Flat inflow scatter plot
flat_mean_model <- lm(year ~ avg_mean_flow, data = flat_average_mean)
flat_r2 <- summary(flat_mean_model)$r.squared

ggplot(data = flat_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Inflow", x = "Year", y = "Mean Inflow (cfs)") +
    annotate("text", x = 1975, y = 3542400, 
           label = paste("R² =", round(flat_r2, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Flat volume line graph
ggplot(data = flat_volume_mean, aes(x = year, y = avg_vol)) +
  geom_line() +
  geom_smooth(method = "lm") +
  annotate("text", x = 1950, y = 630000000, label = "-6.07 * 10^4 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Average Flat Volume", x = "Year", y = "Mean Volume (cf)") +
    theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Flat volume scatter plot
flat_vol_model <- lm(year ~ avg_vol, data = flat_volume_mean)
flat_r3 <- summary(flat_vol_model)$r.squared

ggplot(data = flat_volume_mean, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Average Flat Volume", x = "Year", y = "Mean Volume (cf)") +
  annotate("text", x = 1975, y = 3542400, 
           label = paste("R² =", round(flat_r3, 2)), 
           size = 5) +
   stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)
```

Now we will look at only years 1981 and beyond to capture recent trends.
```{r Flat 1981 and beyond, echo=TRUE}
#Filter out the data to only include 1981 and beyond.
flat_1981 <- flat_inflows %>% 
  filter(year >= 1981)
#Filter out the average median inflow data to only include 1981 and beyond.
flat_1981_median <- flat_average_median %>% 
  filter(year >= 1981)
#Filter out the average mean inflow data to only include 1981 and beyond.
flat_1981_mean <- flat_average_mean %>% 
  filter(year >= 1981)
#Filter out the average volume data to only include 1981 and beyond.
flat_1981_volume <- flat_volume_mean %>% 
  filter(year >= 1981)

#Median Flat inflow line graph
ggplot(data = flat_1981_median, aes(x = year, y = avg_median_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) +
  labs(title = "Median Flat Inflow Since 1981", x = "Year", y = "Median Inflow (cfs)")

#Median Flat inflow scatter plot
flat_med_model_1981 <- lm(year ~ avg_median_flow, data = flat_1981_median)
flat_r4 <- summary(flat_med_model_1981)$r.squared

ggplot(data = flat_1981_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Flat Inflow Since 1981", x = "Year", y = "Median Inflow (cfs)") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(flat_r4, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Flat inflow line graph
ggplot(data = flat_1981_mean, aes(x = year, y = avg_mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
    annotate("text", x = 1990, y = 22000000, label = "4.09 * 10^3 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Mean Flat Inflow Since 1981", x = "Year", y = "Mean Inflow (cfs)") +
      theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Flat inflow scatter plot
flat_mean_model_1981 <- lm(year ~ avg_mean_flow, data = flat_1981_mean)
flat_r5 <- summary(flat_mean_model_1981)$r.squared

ggplot(data = flat_1981_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Inflow Since 1981", x = "Year", y = "Mean Inflow (cfs)") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(flat_r5, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Flat volume line graph
ggplot(data = flat_1981_volume, aes(x = year, y = avg_vol)) +
  geom_line() +
  geom_smooth(method = "lm") +
      annotate("text", x = 1990, y = 630000000, label = "1.06 * 10^5 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Average Flat Volume Since 1981", x = "Year", y = "Mean Volume (cf") +
        theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Flat scatter plot
flat_vol_model_1981 <- lm(year ~ avg_vol, data = flat_1981_volume)
flat_r6 <- summary(flat_vol_model_1981)$r.squared

ggplot(data = flat_1981_volume, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Average Flat Volume Since 1981", x = "Year", y = "Mean Volume (cf") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(flat_r6, 2)), 
           size = 5) +
   stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)
```

Finally, we will look at Neuse outflows. 
```{r Neuse all, echo=TRUE}
#Median Neuse outflow line graph
ggplot(data = neuse_average_median, aes(x = year, y = avg_median_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) +
  labs(title = "Median Neuse outflow", x = "Year", y = "Median Outflow (cfs)")

#Median Neuse outflow scatter plot
neuse_med_model <- lm(year ~ avg_median_flow, data = neuse_average_median)
neuse_r <- summary(neuse_med_model)$r.squared

ggplot(data = neuse_average_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Neuse Outflow", x = "Year", y = "Median Outflow (cfs))") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(neuse_r, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean Neuse outflow line graph
ggplot(data = neuse_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
        annotate("text", x = 2010, y = 93000000, label = "-3.6 * 10^5 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Mean Neuse outflow", x = "Year", y = "Mean Outflow (cfs)") +
      theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Neuse outflow scatter plot
neuse_mean_model <- lm(year ~ avg_mean_flow, data = neuse_average_mean)
neuse_r2 <- summary(neuse_mean_model)$r.squared

ggplot(data = neuse_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Neuse Outflow", x = "Year", y = "Mean Outflow (cfs))") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(neuse_r, 2)), 
           size = 5) +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean volume line graph
ggplot(data = neuse_volume_mean, aes(x = year, y = avg_vol)) +
  geom_line() +
  geom_smooth(method = "lm") +
 annotate("text", x = 2012, y = 2900000000, label = "-1.09 * 10^7 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Average Neuse Volume", x = "Year", y = "Mean Volume (cf)") +
  theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))
  

#Mean volume scatter plot
neuse_vol_model <- lm(year ~ avg_vol, data = neuse_volume_mean)
neuse_r3 <- summary(neuse_vol_model)$r.squared

ggplot(data = neuse_volume_mean, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Average Neuse Volume", x = "Year", y = "Mean Volume (cf)") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(neuse_r3, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) 
```

And now with only 1981 and beyond:
```{r Neuse 1981 and beyond, echo=TRUE}
#Filter out the data to only include 1981 and beyond.
neuse_1981 <- neuse_outflows %>% 
  filter(year >= 1981)
#Filter out the average median outflow data to only include 1981 and beyond.
neuse_1981_median <- neuse_average_median %>% 
  filter(year >= 1981)
#Filter out the average mean outflow data to only include 1981 and beyond.
neuse_1981_mean <- neuse_average_mean %>% 
  filter(year >= 1981)
#Filter out the average volume data to only include 1981 and beyond.
neuse_1981_volume <- neuse_volume_mean %>% 
  filter(year >= 1981)

#Median Neuse outflow line graph
ggplot(data = neuse_1981_median, aes(x = year, y = avg_median_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) +
  labs(title = "Median Neuse Outflow Since 1981", x = "Year", y = "Median Outflow (cfs)")

#Median Neuse outflow scatter plot
neuse_med_model_1981 <- lm(year ~ avg_median_flow, data = neuse_1981_median)
neuse_r4 <- summary(neuse_med_model_1981)$r.squared

ggplot(data = neuse_1981_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Neuse Outflow", x = "Year", y = "Median Outflow (cfs)") +
  annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(neuse_r4, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE) 

#Mean Neuse outflow line graph
ggplot(data = neuse_1981_mean, aes(x = year, y = avg_mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
   annotate("text", x = 2012, y = 81000000, label = "-1.79 * 10^5 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Mean Neuse Outflow Since 1981", x = "Year", y = "Mean Outflow (cfs)") +
    theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))

#Mean Neuse outflow scatter plot
neuse_mean_model_1981 <- lm(year ~ avg_mean_flow, data = neuse_1981_mean)
neuse_r5 <- summary(neuse_mean_model_1981)$r.squared

ggplot(data = neuse_1981_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Neuse Outflow", x = "Year", y = "Mean Outflow (cfs)") +
   annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(neuse_r5, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)

#Mean volume line graph
ggplot(data = neuse_1981_volume, aes(x = year, y = avg_vol)) +
  geom_line() +
  geom_smooth(method = "lm") +
  annotate("text", x = 2012, y = 2800000000, label = "-5.36 * 10^6 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Average Neuse Volume since 1981", x = "Year", y = "Mean Volume (cf)") +
    theme_minimal() +
  theme(text = element_text(family = "Times New Roman"))


#Mean volume scatter plot
neuse_vol_model_1981 <- lm(year ~ avg_vol, data = neuse_1981_volume)
neuse_r6 <- summary(neuse_vol_model_1981)$r.squared

ggplot(data = neuse_1981_volume, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Average Neuse Volume since 1981", x = "Year", y = "Mean Volume (cf)") +
   annotate("text", x = 2010, y = 3542400, 
           label = paste("R² =", round(neuse_r6, 2)), 
           size = 5) +
    stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               parse = TRUE)
```

```{r} 
#Look at combined inflow data
library(readxl)
#Note that it was easier for me to create the combined and projected data in 
#Excel using the FORECAST() function, so I copied the data over into a new sheet, 
#projected, and then loaded that in below.
combined <- read_xlsx("./Data Raw/combined_mean.xlsx")

#Combined average inflows line graph
ggplot(data = combined, aes(x = year, y = combined_mean)) +
  geom_line() +
  geom_smooth(method = "lm") +
  geom_line(aes(color = year >= 2023), size = 0.8) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +  
  scale_size_manual(values = c("FALSE" = 2, "TRUE" = 5)) +
  geom_vline(xintercept = 2023, linetype = "dashed", color = "red") + 
  theme_minimal() +
  annotate("text", x = 1960, y = 30000000, label = "-1.11 * 10^4 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Combined Average Inflows and Projected", x = "Year", y = "Average Inflow (cfs)") +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Times New Roman"))

#Combined average inflows line graph since 1981
ggplot(data = combined, aes(x = year, y = recent_mean)) +
  geom_line() +
  geom_smooth(method = "lm") +
  xlim(1981, max(combined$recent_mean)) +
  geom_line(aes(color = year >= 2023), size = 0.8) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +  
  scale_size_manual(values = c("FALSE" = 2, "TRUE" = 5)) +
  geom_vline(xintercept = 2023, linetype = "dashed", color = "red") + 
  theme_minimal() +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               size = 2.5,
               parse = TRUE,
               hjust = -2, vjust = 11, 
               family = "Times New Roman") +
  labs(title = "Combined Average Inflows Since 1981 and Projected", x = "Year", y = "Average Inflow (cfs)") +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Times New Roman"))
  
#Combined average volumes line graph
ggplot(data = combined, aes(x = year, y = combined_volume)) +
  geom_line() +
  geom_smooth(method = "lm") +
  geom_line(aes(color = year >= 2023), size = 0.8) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +  
  scale_size_manual(values = c("FALSE" = 2, "TRUE" = 5)) +
  geom_vline(xintercept = 2023, linetype = "dashed", color = "red") + 
  theme_minimal() +
    annotate("text", x = 1975, y = 900000000, label = "-3.46 * 10^5 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Combined Inflow Volumes and Projected", x = "Year", y = "Volume (cf)") +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Times New Roman"))



#Combined average volumes since 1981 line graph
ggplot(data = combined, aes(x = year, y = recent_volume)) +
  geom_line() +
  geom_smooth(method = "lm") +
  geom_line(aes(color = year >= 2023), size = 0.8) +
  xlim(1981, max(combined$recent_volume)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +  
  scale_size_manual(values = c("FALSE" = 2, "TRUE" = 5)) +
  geom_vline(xintercept = 2023, linetype = "dashed", color = "red") + 
  theme_minimal() +
  stat_poly_eq(aes(label = after_stat(eq.label)), 
               formula = y ~ x, 
               size = 2.5,
               parse = TRUE,
               hjust = -1.5, vjust = 5,
               family = "Times New Roman") +
  labs(title = "Combined Inflow Volumes Since 1981 and Projected", x = "Year", y = "Volume (cf)") +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Times New Roman"))


#Neuse average outflow projected
#Note that it was easier for me to create the projected data in Excel using the
#FORECAST() function, so I copied the data over into a new sheet, projected, and
#then loaded that in below.
neuse <- read_xlsx("./Data Raw/neuse_projected.xlsx")

ggplot(data = neuse, aes(x = year, y = mean_flow)) +
  geom_line() +
  geom_smooth(method = "lm") +
  geom_line(aes(color = year >= 2023), size = 0.8) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +  
  scale_size_manual(values = c("FALSE" = 2, "TRUE" = 5)) +
  geom_vline(xintercept = 2023, linetype = "dashed", color = "red") + 
  theme_minimal() +
  annotate("text", x = 2034, y = 66000000, label = "-3.46 * 10^5 * x cfs/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Neuse Mean and Projected Outflows", x = "Year", y = "Mean Outflow (cfs)") +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Times New Roman"))


#Neuse volume projected
ggplot(data = neuse, aes(x = year, y = volume)) +
  geom_line() +
  geom_smooth(method = "lm") +
  geom_line(aes(color = year >= 2023), size = 0.8) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +  
  scale_size_manual(values = c("FALSE" = 2, "TRUE" = 5)) +
  geom_vline(xintercept = 2023, linetype = "dashed", color = "red") + 
  theme_minimal() +
  annotate("text", x = 2034, y = 2000000000, label = "-1.09 * 10^7 * x cf/year", size = 2.5,
           family = "Times New Roman") +
  labs(title = "Projected Neuse Outflow Volume", x = "Year", y = "Volume (cf)") +
  theme(legend.position = "none") +
  theme(text = element_text(family = "Times New Roman"))
```
