---
title: "Assignment 1: Aidan Power, Fiona Price, and Emily Burghart"
output: html_document
date: "2024-09-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the packages and data

Loading in the packages:

```{r setup}
library(tidyverse)
library(ggplot2)
library(here)
library(dplyr)
```

Now, we will load in the Eno data and make separate dataframes for the median 
outflow, mean outflow, and volume.

```{r Eno data}
#Eno data:
eno_outflows <- read.csv(
  file = here("./Data Raw/Eno_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median outflow:
eno_average_median <- eno_outflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean outflow:
eno_average_mean <- eno_outflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
eno_volume_mean <- eno_outflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

Now, we will load in the Flat data and make separate dataframes for the median 
outflow, mean outflow, and volume.

```{r Flats data}
#Flats data:
flat_outflows <- read.csv(
  file = here("./Data Raw/Flat_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median outflow:
flat_average_median <- flat_outflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean outflow:
flat_average_mean <- flat_outflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
flat_volume_mean <- flat_outflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

Now, we will load in the Neuse data and make separate dataframes for the median 
inflow, mean inflow, and volume.

```{r Neuse data}
#Neuse data:
neuse_inflows <- read.csv(
  file = here("./Data Raw/Falls_Lake_outlet_Neuse_River_monthly_flow_data_through-2022.csv"),
  stringsAsFactors = TRUE)
#Average of median inflow:
neuse_average_median <- neuse_inflows %>% 
  group_by(year) %>% 
  summarise(avg_median_flow = mean(median_flow, na.rm = TRUE))
#Average of mean inflow:
neuse_average_mean <- neuse_inflows %>% 
  group_by(year) %>% 
  summarise(avg_mean_flow = mean(mean_flow, na.rm = TRUE))
#Average of volume:
neuse_volume_mean <- neuse_inflows %>% 
  group_by(year) %>% 
  summarise(avg_vol = mean(total_volume, na.rm = TRUE))
```

## Plots
Now, we will start making graphs. First, we will look at Eno. We will start with
all years.

```{r Eno all, echo=FALSE}
#Median Eno outflow
ggplot(data = (eno_outflows)) +
  geom_col(data = eno_outflows, aes(x = year, y = median_flow)) +
  labs(title = "Median Eno Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")
ggplot(data = eno_average_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Eno Outflow", x = "Year", y = "Median Inflow (in Acre-Feet)")

#Mean Eno outflow
ggplot(data = eno_outflows) +
  geom_col(data = eno_outflows, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Eno Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")
ggplot(data = eno_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Outflow", x = "Year", y = "Mean Inflow (in Acre-Feet)")

#Mean volume
ggplot(data = eno_volume_mean) +
  geom_col(data = eno_volume_mean, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Eno Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = eno_volume_mean, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```

There is a gap in Eno data from 1971-1985. Now, we will look only at the post-1985
data.

```{r Eno 1985 and beyond, echo=FALSE}
#Filter out the data to only include 1985 and beyond.
eno_filtered <- eno_outflows %>% 
  filter(year >= 1985)
#Filter out the average median outflow data to only include 1985 and beyond.
eno_avemedian_filtered <- eno_average_median %>% 
  filter(year >= 1985)
#Filter out the average mean outflow data to only include 1985 and beyond.
eno_avemean_filtered <- eno_average_mean %>% 
  filter(year >= 1985)
#Filter out the average volume data to only include 1985 and beyond.
eno_vol_filtered <- eno_volume_mean %>% 
  filter(year >= 1985)

#Median Eno outflow
ggplot(data = (eno_filtered)) +
  geom_col(data = eno_filtered, aes(x = year, y = median_flow)) +
  labs(title = "Median Eno Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")
ggplot(data = eno_avemedian_filtered, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Eno Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")

#Mean Eno outflow
ggplot(data = eno_filtered) +
  geom_col(data = eno_filtered, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Eno Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")
ggplot(data = eno_avemean_filtered, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Outflow", x = "Year", y = "Mean Inflow (in Acre-Feet)")

#Mean Eno volume
ggplot(data = eno_vol_filtered) +
  geom_col(data = eno_vol_filtered, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Eno Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = eno_vol_filtered, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```

Lastly, we will look at only years 2000 and beyond to capture recent trends.
```{r Eno 2000 and beyond, echo=FALSE}
#Filter out the data to only include 2000 and beyond.
eno_2000 <- eno_outflows %>% 
  filter(year >= 2000)
#Filter out the average median outflow data to only include 2000 and beyond.
eno_2000_median <- eno_average_median %>% 
  filter(year >= 2000)
#Filter out the average mean outflow data to only include 2000 and beyond.
eno_2000_mean <- eno_average_mean %>% 
  filter(year >= 2000)
#Filter out the average volume data to only include 2000 and beyond.
eno_2000_volume <- eno_volume_mean %>% 
  filter(year >= 2000)

#Median Eno outflow
ggplot(data = eno_2000) +
  geom_col(data = eno_2000, aes(x = year, y = median_flow)) +
  labs(title = "Median Flat Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")
ggplot(data = eno_2000_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Flat Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")

#Mean Eno outflow
ggplot(data = eno_2000) +
  geom_col(data = eno_2000, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Flat Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")
ggplot(data = eno_2000_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")

#Mean Eno volume
ggplot(data = eno_2000_volume) +
  geom_col(data = eno_2000_volume, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Eno Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = eno_2000_volume, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Eno Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```

Now, we will look at Flat data.
```{r Flat all, echo=FALSE}
#Median Flat outflow
ggplot(data = (flat_outflows)) +
  geom_col(data = flat_outflows, aes(x = year, y = median_flow)) +
  labs(title = "Median Flat Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")
ggplot(data = flat_average_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Flat Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")

#Mean Flat outflow
ggplot(data = (flat_outflows)) +
  geom_col(data = flat_outflows, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Flat Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")
ggplot(data = flat_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")

#Mean volume
ggplot(data = flat_volume_mean) +
  geom_col(data = flat_volume_mean, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Flat Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = flat_volume_mean, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```

Now we will look at only years 2000 and beyond to capture recent trends.
```{r Flat 2000 and beyond, echo=FALSE}
#Filter out the data to only include 2000 and beyond.
flat_2000 <- flat_outflows %>% 
  filter(year >= 2000)
#Filter out the average median outflow data to only include 2000 and beyond.
flat_2000_median <- flat_average_median %>% 
  filter(year >= 2000)
#Filter out the average mean outflow data to only include 2000 and beyond.
flat_2000_mean <- flat_average_mean %>% 
  filter(year >= 2000)
#Filter out the average volume data to only include 2000 and beyond.
flat_2000_volume <- flat_volume_mean %>% 
  filter(year >= 2000)

#Median Flat outflow
ggplot(data = flat_2000) +
  geom_col(data = flat_2000, aes(x = year, y = median_flow)) +
  labs(title = "Median Flat Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")
ggplot(data = flat_2000_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Flat Outflow", x = "Year", y = "Median Outflow (in Acre-Feet)")

#Mean Flat outflow
ggplot(data = flat_2000) +
  geom_col(data = flat_2000, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Flat Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")
ggplot(data = flat_2000_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Outflow", x = "Year", y = "Mean Outflow (in Acre-Feet)")

#Mean Flat volume
ggplot(data = flat_2000_volume) +
  geom_col(data = flat_2000_volume, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Flat Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = flat_2000_volume, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Flat Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```

Finally, we will look at Neuse inflows.
```{r Neuse all, echo=FALSE}
#Median Neuse inflow
ggplot(data = (neuse_inflows)) +
  geom_col(data = neuse_inflows, aes(x = year, y = median_flow)) +
  labs(title = "Median Neuse Inflow", x = "Year", y = "Median Inflow (in Acre-Feet)")
ggplot(data = neuse_average_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Neuse Inflow", x = "Year", y = "Median Inflow (in Acre-Feet)")

#Mean Neuse inflow
ggplot(data = neuse_inflows) +
  geom_col(data = neuse_inflows, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Neuse Inflow", x = "Year", y = "Mean Inflow (in Acre-Feet)")
ggplot(data = neuse_average_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Neuse Inflow", x = "Year", y = "Mean Inflow (in Acre-Feet)")

#Mean volume
ggplot(data = neuse_volume_mean) +
  geom_col(data = neuse_volume_mean, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Neuse Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = neuse_volume_mean, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Neuse Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```

And now with only 2000 and beyond:
```{r Neuse 2000 and beyond, echo=FALSE}
#Filter out the data to only include 2000 and beyond.
neuse_2000 <- neuse_inflows %>% 
  filter(year >= 2000)
#Filter out the average median outflow data to only include 2000 and beyond.
neuse_2000_median <- neuse_average_median %>% 
  filter(year >= 2000)
#Filter out the average mean outflow data to only include 2000 and beyond.
neuse_2000_mean <- neuse_average_mean %>% 
  filter(year >= 2000)
#Filter out the average volume data to only include 2000 and beyond.
neuse_2000_volume <- neuse_volume_mean %>% 
  filter(year >= 2000)

#Median Neuse inflow
ggplot(data = (neuse_2000)) +
  geom_col(data = neuse_2000, aes(x = year, y = median_flow)) +
  labs(title = "Median Neuse Inflow", x = "Year", y = "Median Inflow (in Acre-Feet)")
ggplot(data = neuse_2000_median, aes(x = year, y = avg_median_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Median Neuse Inflow", x = "Year", y = "Median Inflow (in Acre-Feet)")

#Mean Neuse inflow
ggplot(data = neuse_2000) +
  geom_col(data = neuse_2000, aes(x = year, y = mean_flow)) +
  labs(title = "Mean Neuse Inflow", x = "Year", y = "Mean Inflow (in Acre-Feet)")
ggplot(data = neuse_2000_mean, aes(x = year, y = avg_mean_flow)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Neuse Inflow", x = "Year", y = "Mean Inflow (in Acre-Feet)")

#Mean Neuse volume
ggplot(data = neuse_2000_volume) +
  geom_col(data = neuse_2000_volume, aes(x = year, y = avg_vol)) +
  labs(title = "Mean Neuse Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
ggplot(data = neuse_2000_volume, aes(x = year, y = avg_vol)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Mean Neuse Volume", x = "Year", y = "Mean Volume (in Acre-Feet)")
```